---
title: "H1B Visa Petitions Analysis"
author: "Ramya Bhat"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##H1B Visa

There is a special kind of visa for non-immigrants of United States who are intersted in working for companies situated in USA. The H-1B is an employment-based, non-immigrant visa category for temporary foreign workers in the United States. For a foreign national to apply for H1-B visa, an US employer must offer a job and petition for H-1B visa with the US immigration department. This is the most common visa status applied for and held by international students once they complete college/ higher education (Masters, PhD) and work in a full-time position. Under the H1B visa, any company can employ a foreign worker for up to six years. Filing of a H1B visa is not in hands of the individuals, only the employer is allowed to file the petition for the respective employee.


```{r echo=FALSE,message=FALSE,warning=FALSE}
#-------libraries and data read-------#
library(sampling)
library(plotly)
library(stringr)
H1B_Dataset <- read.csv("/Users/ramyabhat/Desktop/ram-544/project/1. Master H1B Dataset.csv")
```

##Dataset Description

The H-1B Dataset selected for this project contains data from employer’s Labor Condition
Application and the case certification determinations processed by the Office of Foreign
Labor Certification (OFLC). The dataset has more than 500,000 records overall. The relevant coloumns included in our data set are: employer name,job title,case number,visa class, employment start and end dates,case status and so on.


##Goal of the project

The objective of this project is to have a deeper insight about the H1B applications filed in United States of America and also to predict the case status of an application submitted by the employer to hire non-immigrant workers under the H-1B visa program.


##Data Preprocessing

Data preprocesing involves transforming data into a basic form that makes it easy to work with. Our set of data containing more than 500,000 rows must be pre-processed and filtered down to suit the requirements of our project. 

The challenge we all face in data preprocessing is to handle null, N/A vaues and the outliers. All the rows of the class attributes that has N/A or null values in them must been removed and outliers has to be dealt with.Activities done in this step includes detecting the presence of missing (NA) values and outliers, or duplicate data.


```{r echo=FALSE,message=FALSE,warning=FALSE}
#-------Data pre-processing-------# 
H1B_Dataset <- H1B_Dataset[H1B_Dataset$VISA_CLASS=='H1B' & H1B_Dataset$DECISION_YEAR=='2017' & H1B_Dataset$PREVAILING_WAGE !='0' & H1B_Dataset$EMPLOYER_COUNTRY=="UNITED STATES OF AMERICA",]

Dataset_summary <- summary(H1B_Dataset$PREVAILING_WAGE)
Dataset_summary

q1 <-(Dataset_summary[2]-(1.5 * (Dataset_summary[5]-Dataset_summary[2]))) 
q3 <- (Dataset_summary[5]+(1.5 * (Dataset_summary[5]-Dataset_summary[2])))

index_outlrs <- which(H1B_Dataset$PREVAILING_WAGE<q1 | H1B_Dataset$PREVAILING_WAGE>=q3)

H1B_Dataset <- H1B_Dataset[-index_outlrs,]

```

##Categorizing the H1B applicants into groups

```{r echo=FALSE,message=FALSE,warning=FALSE}
#-------Piechart-------#
plot_ly(H1B_Dataset, labels = ~names(table(H1B_Dataset$CASE_STATUS)), 
        values = ~table(H1B_Dataset$CASE_STATUS), type = 'pie',
        textposition = 'inside',textinfo = 'label+percent') %>% layout(title = "Status of Applications",
                                                                       xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE), yaxis = list(showgrid = FALSE,
                                                                                                                                                              zeroline = FALSE, showticklabels = FALSE)) 


```


## Determining the distribution of salary of the applicants based on the status of their job.

Below is a box plot and histogram of the prevailing wage distribution for each group of H1B applicants.

```{r echo=FALSE,message=FALSE,warning=FALSE}
#----boxplots and histograms----#
First_plot <- H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "CERTIFIED"]
Sal_plot<- plot_ly(x = ~H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "CERTIFIED"], type = "box" , name = "CERTIFIED") %>% 
  add_trace(x = ~H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "DENIED"], name = "DENIED") %>%
  add_trace(x = ~H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "CERTIFIEDWITHDRAWN"], name = "CERTIFIEDWITHDRAWN") %>%
  add_trace(x = ~H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "WITHDRAWN"], name = "WITHDRAWN") %>% layout(title = "Salary distribution of Groups",
                                                                                                                 xaxis = list(title = "Prevailing wage", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE), yaxis = list(showgrid = FALSE,
                                                                                                                                                                                                                                   zeroline = FALSE, showticklabels = FALSE))
Sal_plot
```
                                                .

```{r echo=FALSE,message=FALSE,warning=FALSE}
#----histograms----#
histogram_1 <- plot_ly(x=H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "CERTIFIED"], type='histogram' ,nbinsx=100, color = "", colors = c("Green"), name="Certified")
histogram_2 <- plot_ly(x=H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "DENIED"], type='histogram' ,nbinsx=100, color = "", colors = c("Red"), name="Denied")
histogram_3 <- plot_ly(x=H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "CERTIFIEDWITHDRAWN"], type='histogram' ,nbinsx=100, color = "", colors = c("Orange"), name="Certified-Withdrawn")
histogram_4 <- plot_ly(x=H1B_Dataset$PREVAILING_WAGE[H1B_Dataset$CASE_STATUS == "WITHDRAWN"], type='histogram' ,nbinsx=100, color = "", colors = c("Blue"), name="Withdrawn")

Dist_Sal <- subplot(nrows=2,histogram_1,histogram_2,histogram_3,histogram_4) %>% layout(title = "Distribution of Salaries ")
Dist_Sal

```

## Distribution of applicants by measure of Salary. 

```{r echo=FALSE,message=FALSE,warning=FALSE}
 #Histogram For Salary
plot_ly(x=H1B_Dataset$PREVAILING_WAGE, type='histogram' ,nbinsx=100, color = "", colors = c("cyan"), name="10")

```


##Central Limit Theorem

The central limit theorem states that the distribution of sample means, taken from independent random sample sizes, follows a normal distribution even if the original population is not normally distributed. As a result we can apply statistical techniques that assume normality even when the population is non normal. 
In this data set the applicability of the central limit theorem can be shown by using the Prevailing_wage attribute . As we can see from the above histogram, the salary distribution of all applicants have a normal distribution hence Prevailing_wage is used here for central limit theorem. Below are histograms showing the sample means of 1000 random samples of sample size 10, 20, 30, and 40 following a normal distribution.

```{r echo=FALSE,message=FALSE,warning=FALSE}
#-------Central Limit Theorem-------#
Sum <- as.numeric(H1B_Dataset$PREVAILING_WAGE)

samples <- 10000

mean_size_10 <- numeric(samples)
mean_size_20 <- numeric(samples)
mean_size_30 <- numeric(samples)
mean_size_40 <- numeric(samples)

set.seed(123)

for(i in 1:samples){
  mean_size_10[i]<- mean(sample(Sum, size=10, replace = T))
  mean_size_20[i]<- mean(sample(Sum, size=20, replace = T))
  mean_size_30[i]<- mean(sample(Sum, size=30, replace = T))
  mean_size_40[i]<- mean(sample(Sum, size=40, replace = T))
}


plot_1 <- plot_ly(x=mean_size_10, type='histogram' ,nbinsx=100, color = "", colors = c("Yellow"), name="10")

plot_2 <- plot_ly(x=mean_size_20, type='histogram' ,nbinsx=100, color = "", colors = c("brown"), name="20") 

plot_3 <- plot_ly(x=mean_size_30, type='histogram' ,nbinsx=100, color = "", colors = c("navy"),name="30")

plot_4 <- plot_ly(x=mean_size_40, type='histogram' ,nbinsx=100, color = "", colors = c("Aquamarine"), name="40")

cat("Population , Mean =",round(mean(Sum),3), ", SD = ", round(sd(Sum),3))
cat("Sampe Size 10, Mean =", round(mean(mean_size_10),3), ', SD =',round(sd(Sum)/(sqrt(10)),3),"\nSampe Size 20, Mean =", round(mean(mean_size_20),3),", SD =",round(sd(Sum)/(sqrt(20)),3),"\nSampe Size 30, Mean =", round(mean(mean_size_30),3),", SD =",round(sd(Sum)/(sqrt(30)),3),"\nSampe Size 40, Mean =", round(mean(mean_size_40),3),", SD =",round(sd(Sum)/(sqrt(40)),3))
subplot(nrows=2,plot_1,plot_2,plot_3,plot_4) %>% layout(title = "Sampling Distribution ")
```

##Sampling

Sampling is a technique to select a representative portion of the population to perform a study on. There are many different sampling techniques ->
1) Simple random sampling - a basic sampling technique where individual subjects are selected from a larger group. In this case, every sample has the same chance of getting picked. 
2) Systematic sampling - Systematic sampling is a method where samples are selected via a fixed periodic interval. The interval is calculated by dividing the whole population sample by the desired sample size.
3) Stratified sampling. 

For this project - Simple random sampling with replacement, simple random sampling without replacement, systematic sampling, and unequal probability sampling has been be used as sampling methods.

```{r echo=FALSE,message=FALSE,warning=FALSE}

#---Simple random sampling with replacement---#
set.seed(123)

samp_wrep <- srswr(100, nrow(H1B_Dataset))

select_data <- (1:nrow(H1B_Dataset))[samp_wrep!=0]
select_data <- rep(select_data,samp_wrep[samp_wrep!=0])
sample_srswr <- H1B_Dataset[select_data,] 

orig_data <- plot_ly(x=H1B_Dataset$PREVAILING_WAGE, type='histogram' ,nbinsx=100, color = "", colors = c("Greys"), name="Salary")
histo_srswr <- plot_ly(x=sample_srswr$PREVAILING_WAGE, type='histogram' ,nbinsx=100, color = "", colors = c("Blues"), name="SRSWR")
histo_srswr

#---Simple random sampling without replacement---#
set.seed(123)

samp_worep <- srswor(100, nrow(H1B_Dataset))
sample_srswor <- H1B_Dataset[samp_worep!=0,]

histo_srswor <- plot_ly(x=sample_srswor$PREVAILING_WAGE, type='histogram' ,nbinsx=100, color = "", colors = c("Reds"), name="SRSWOR")
histo_srswor

#---Systematic Sampling---#
set.seed(123)

N <- nrow(H1B_Dataset)
n <- 100
k <- ceiling(N / n)
r <- sample(k, 1)
s <- seq(r, by = k, length = n)
sample_syst <- H1B_Dataset[s, ]

histo_sys <- plot_ly(x=sample_syst$PREVAILING_WAGE, type='histogram' ,nbinsx=100, color = "", colors = c("BLUE"), name="Systematic")
histo_sys


#---unequal prob systematic sampling---#
set.seed(123)

unequal_prob <- inclusionprobabilities(H1B_Dataset$PREVAILING_WAGE, 100)

s <- UPsystematic(unequal_prob)

sample.unpsyst <- H1B_Dataset[s != 0, ]


histo_unpsyst <- plot_ly(x=sample.unpsyst$PREVAILING_WAGE, type='histogram' ,nbinsx=100, color = "", colors = c("Aquamarine"), name="Unequal Prob Systematic")
histo_unpsyst
```

##Confidence Intervals

```{r echo=FALSE,message=FALSE,warning=FALSE}
#Population mean, sd, confidence interval and aplha level:-
#-------Confidence-------#
dataset.mean <- mean(H1B_Dataset$PREVAILING_WAGE)
dataset.sd <- sd(H1B_Dataset$PREVAILING_WAGE)
confid <- c(80, 90)
c_alpha <- 1 - confid/100

cat("Prevailing wage : mean =",dataset.mean," and sd =",dataset.sd)


for (i in c_alpha) {
  str_srswr <- sprintf("%2d%% Conf Level (c_alpha = %.2f), CI = %.2f - %.2f",
                       100*(1-i), i, 
                       dataset.mean - qnorm(1-i/2) * dataset.sd,
                       dataset.mean + qnorm(1-i/2) * dataset.sd)
  cat(str_srswr,"\n")
}

#srswr#
sd.sample.means_srswr <- dataset.sd/sqrt(nrow(sample_srswr))
xbar_srswr <- mean(sample_srswr$PREVAILING_WAGE)
cat("SRSWR : mean =",xbar_srswr," and sd =",sd.sample.means_srswr)

for (i in c_alpha) {
  str_srswr <- sprintf("%2d%% Conf Level (c_alpha = %.2f), CI = %.2f - %.2f",
                       100*(1-i), i, 
                       xbar_srswr - qnorm(1-i/2) * sd.sample.means_srswr,
                       xbar_srswr + qnorm(1-i/2) * sd.sample.means_srswr)
  cat(str_srswr,"\n")
}

#srswor#
sd.sample.means_srswor <- dataset.sd/sqrt(nrow(sample_srswor))
xbar_srswor <- mean(sample_srswor$PREVAILING_WAGE)
cat("SRSWOR : mean =",xbar_srswor," and sd =",sd.sample.means_srswor)

for (i in c_alpha) {
  str_srswor <- sprintf("%2d%% Conf Level (c_alpha = %.2f), CI = %.2f - %.2f",
                        100*(1-i), i, 
                        xbar_srswor - qnorm(1-i/2) * sd.sample.means_srswor,
                        xbar_srswor + qnorm(1-i/2) * sd.sample.means_srswor)
  cat(str_srswor,"\n")
}


#systematic#
sd.sample.means_syst <- dataset.sd/sqrt(nrow(sample_syst))
xbar_syst <- mean(sample_syst$PREVAILING_WAGE[!is.na(sample_syst$PREVAILING_WAGE)])
cat("SRSWOR : mean =",xbar_syst," and sd =",sd.sample.means_syst)

for (i in c_alpha) {
  str_syst <- sprintf("%2d%% Conf Level (c_alpha = %.2f), CI = %.2f - %.2f",
                            100*(1-i), i, 
                            xbar_syst - qnorm(1-i/2) * sd.sample.means_syst,
                            xbar_syst + qnorm(1-i/2) * sd.sample.means_syst)
  cat(str_syst,"\n")
}

#Upsystematic#
sd.sample.means_upsyst <- dataset.sd/sqrt(nrow(sample.unpsyst))
xbar_upsyst <- mean(sample.unpsyst$PREVAILING_WAGE[!is.na(sample.unpsyst$PREVAILING_WAGE)])
cat("UPSystematic : mean =",xbar_upsyst," and sd =",sd.sample.means_upsyst)

for (i in c_alpha) {
  str_upsyst <- sprintf("%2d%% Conf Level (c_alpha = %.2f), CI = %.2f - %.2f",
                              100*(1-i), i, 
                              xbar_upsyst - qnorm(1-i/2) * sd.sample.means_upsyst,
                              xbar_upsyst + qnorm(1-i/2) * sd.sample.means_upsyst)
  cat(str_upsyst,"\n")
}
```

##Jobs that are of more demand in market

Below bar chart depicts the top six job designations that opt for H1B visa and their mean salary.

```{r echo=FALSE,message=FALSE,warning=FALSE}
 #Ondemand
dem_2 <- numeric(1)
dem <- names(tail(sort(table(H1B_Dataset$SOC_NAME))))

for (i in 1:length(dem)) {
  
  dem_2[i] <- mean((H1B_Dataset$PREVAILING_WAGE[as.character(H1B_Dataset$SOC_NAME) == dem[i]]))
}


job_freq_plot <- plot_ly(x = ~dem, y = ~dem_2, color = ~dem, type = "bar" ) %>% layout(title = "Popular Jobs with average Income per annum", xaxis = list(title = ""), yaxis = list(title = "Salary"))
job_freq_plot

```

## Analysis of the applications received by state

The geo plot gives us the information about number of H1B visa applicantions received from individual states in USA.

```{r echo=FALSE,message=FALSE,warning=FALSE}
 #statewise
l <- list(color = toRGB("white"), width = 2)

g <- list(scope = 'usa', projection = list(type = 'albers usa'), showlakes = TRUE, lakecolor = toRGB('white'))

m <- list( l = 50, r = 50, b = 100,  t = 100,  pad = 4)

employer_by_state <- plot_geo(H1B_Dataset, locationmode = 'USA-states') %>%
  add_trace(z = as.vector(sort(table(H1B_Dataset$EMPLOYER_STATE))), 
            locations = names(sort(table(H1B_Dataset$EMPLOYER_STATE))), 
            color = as.vector(sort(table(H1B_Dataset$EMPLOYER_STATE))), 
            locations = names(table(H1B_Dataset$STATE)), 
            colors = 'Reds') %>% 
  colorbar(title = "Number of H1B") %>%  
  layout(title = 'H1B Visa Applications By State', geo = g,autosize = F)

employer_by_state
```
                                                                .
```{r include=TRUE}
```
